---
title: "Background resource extension"
output:
  html_document:
    toc: yes
    fig_width: 10
    fig_height: 5
---


## Load libraries

Mizer 2.0.1 updated at 23/05/2020 or later version is needed to use the background extension. 

 First and best option: download the most updated version of the package directly from GitHub using devtools::install_github("sizespectrum/mizer") 


```{r message=FALSE, warning=FALSE}
devtools::install_github("sizespectrum/mizer") # run only once
library(mizer)
library(viridisLite)
library(viridis)
library(ggplot2)
library(ggrepel)
library(tidyverse)
remotes::install_github("sizespectrum/mizerExperimental")
library(mizerExperimental)
library(mizerHowTo)
source("utility.R")
remotes::install_github("sizespectrum/mizerMR")
library(mizerMR)
library(tidyverse)
```

## Set up multispecies model
my parameters

```{r message=FALSE, warning=FALSE}
# loading calibrated simulation
sim <- readRDS("sim_optim3.RDS")
params <- sim@params
names(params@species_params)[7]<-"biomass_observed"
params@interaction[3,]<-0
params@interaction[4,]<-1
#params@species_params$R_max<-Inf
params@species_params$R_max<-params@species_params$w_inf^-1.5
params@species_params$erepro<-0.01
params@species_params$h<-80
# sim <- project(params, t_max = 5)
# plot(sim)


# check if start  from scratch
speciesParams<-params@species_params
inter<-params@interaction
param <- newMultispeciesParams(speciesParams, kappa = 0.1,inter)
sim<- project(param, t_max = 200, effort = 0)
plot(sim)

params<-steady(param)
plotBiomassVsSpecies(params) ##KP: function doesn't work, not sure what package I'm missing
plotBiomassObservedVsModel(params) #KP: function doesn't work

params <- params %>% calibrateBiomass() %>% matchBiomasses() %>% steady() #KP: function doesn't work
plotBiomassVsSpecies(params)

sim<- project(params, t_max = 200, effort = 0)
plot(sim) 

```


Need to edit resource interaction matrix? Default is 1 
```{r message=FALSE, warning=FALSE, eval = FALSE}

# set  up three  background resources here for a coral reef community  model.
# should  be able  to give the reosurces names but didn't work when I tried  it , so just need to keep  track of what's  what. 1= algae,  2 = benthic invertebrates, 3 = pelagic plankton

#could change  size ranges here  but just leaving the  same for the time being
library(tibble)
resource_params <- tribble(
    ~resource,  ~kappa, ~lambda, ~r_pp, ~w_min, ~w_max,
 "Resource 1",    0.1,    1.05,   5,   1e-12 ,   100,
 "Resource 2",    10,    2.05,    5,   1e-12,    100,
  "Resource 3",    0.1,    2.05,   10,   1e-12,    NA,)


resource_interaction <- matrix(runif(5*3), nrow = 5, ncol = 3)
resource_interaction[,1] <-c(0,0,1,0, 0)
resource_interaction[,2] <-c(1,1,0,1,0)
resource_interaction[,3] <-c(0,0,0,1,1)

# We may not need  to incorporate extra density dependence (R_max). Re-set R_max  to  "Inf" in first case.
# species_params(params)$R_max<-1e10
# Error: kappa is not a number (a length one numeric vector).
# Leave as is for now, have checked with  Gustav.

# speciesParams<-params@species_params
# inter<-params@interaction

params <- setMultipleResources(params, resource_params) #KP: function doesn't work
#params <- newMRParams(speciesParams, resource_params,inter)
sim <- project(params, t_max = 200, t_save = 0.2)
# sim2 <- project(params, t_max = 2, t_save = 0.2, initial_n = sim@n[200,,],initial_n_pp = sim@n_pp[200,])
animateSpectra(sim)
plotSpectra(sim)
plotlyBiomass(sim)
plotlyGrowthCurves(params,percentage = F,
  species_panel = T,max_age = 20)

# need  to re-do Romain's diet plots with the mizerMR  structure
# plotDiet2(params,species="HMD")
# plotDiet2(params,species="sessileCorallivore")
# plotDiet2(params,species="piscivore")

# ask Romain to fix plotDiet

plot(sim) #  note: doesnt show resources but still handy to chekc stuff

params@species_params

# check  tuneParams and download  params - doesnt work as ignores  multiple resources
# tuneParams(params)
# tuned<-readRDS("tuneparams.rds")
# sim_tuned <- project(tuned, t_max = 200, t_save = 0.2)
# sim2 <- project(params, t_max = 2, t_save = 0.2, initial_n = 
# plotSpectra(sim_tuned)
```

If we cannot live with the mismatch between model and observation biomasses (they  seem OK considering  not log scale?) below we  could set up  an optimisation at this stage. We could use the whole size spectrum for the MPA to have  an even stronger basis for doing  it. Would  need to create size bins that are the same as mizer's from the data but that should not be a problem.

Or else we could carry out fishing scenarios and look at relative changes.

Next set up fishing...what questions....which scenarios....what outputs?

```{r}
library(mizerExperimental)


## first re-calibrate and get steady state

plotBiomassVsSpecies(params)
plotBiomassObservedVsModel(params,ratio=T)


sim2<-project(params,t_max=500,t_save = 0.2,effort = 0.0)
plotSpectra(sim2,power=0)
plot(sim2)

params <- params %>% calibrateBiomass() %>% matchBiomasses() %>% steady()
plotBiomassVsSpecies(params)

params2 <- params

params2 <- params2 %>% calibrateBiomass() %>% matchBiomasses() %>% steady()
plotBiomassVsSpecies(params2)

sim2<-project(params,t_max=200,t_save = 0.2,effort = 0.0)
plotSpectra(sim2,power=0)
plot(sim2)

### then add fishing 
```


Kirsty Adding Fishing....
```{r}
paramsfishtest <- setFishing(params, catchability = 0.5, comment_catchability=, initial_effort = 0.5 )

```

```{r}
sim2<-project(params2,t_max=200,t_save = 0.2,effort = 0.2)

# need to set up gears 

#animateSpectra(sim2)

plotYieldVsF(params,species = "sessileCorallivore")
plotYieldVsF(params,species = "mobileinvertivore")
plotYieldVsF(params,species = "HMD")
plotYieldVsF(params,species = "piscivore")
plotYieldVsF(params,species = "planktivore")

plotBiomass(sim2)
```
